Boutiques: a flexible framework for automated application integration in computing platforms

### 2016-11-03 paper.tex
First introduction draft

\usepackage{minted} % Used for JSON highlighting
\usepackage{pdfcomment} % Used for notes and todos
\usepackage[margin=1in]{geometry} % Increases margins
\usepackage[table]{xcolor} % Color definitions
\usepackage{algorithm} % Algorithm float
\usepackage{algpseudocode} % Algorithmic environment

@@

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\todo}[2]{\color{red}TODO: #1\color{black}}
\algrenewcommand{\algorithmiccomment}[1]{\# \textit{#1}} % so that comments in algorithms are preprended by '#' instead of a small right arrow

@@

@@

Why Linux. Define application. Define computing platforms. Explain the
problem and why it is difficult. Approach: start simple and enrich the
spec when new applications come. Refer to computing infrastructures
and the need for parallelization. Who are
the developers. Refer to neuroinformatics field.

Vocabulary (check consistency across the paper):
\begin{itemize}
\item tool vs. application.
\item descriptor vs. manifest.
\item task vs. invocation.
\end{itemize}

\section{System description}

@@

@@

\section{Results}

@@

\paragraph{Pegasus}

@@

\paragraph{SPINE}

@@

\paragraph{VIP}

>>>>

\usepackage{minted} % Used for JSON highlighting
\usepackage{pdfcomment} % Used for notes and todos
\usepackage[margin=1in]{geometry} % Increases margins
\usepackage[table]{xcolor} % Color definitions
\usepackage{algorithm} % Algorithm float
\usepackage{algpseudocode} % Algorithmic environment
\usepackage{xspace}

@@

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\todo}[1]{\color{red}TODO: #1\color{black}}
\newcommand{\boutiques}{Boutiques\xspace}
\algrenewcommand{\algorithmiccomment}[1]{\# \textit{#1}} % so that comments in algorithms are preprended by '#' instead of a small right arrow

@@

\abstract{Porting applications to execution platforms is a critical
  step to enable their sharing among scientific communities and their
  efficient execution on computing infrastructures. However,
  application porting remains a costly human effort that consists of
  1)~installing the application on the target infrastructure,
  2)~describing the application in a format compatible with the
  execution platform, and 3)~generating proper user interfaces. Due to
  the variety of platforms, application porting efforts are often
  replicated several times while mutualization would save cost and
  improve the quality of the ported applications. We describe
  \emph{boutiques}, a system that allows for automatic import and
  exchange of applications across platforms. Boutiques relies on Linux
  containers to solve the problem of application installation in a
  lightweight manner. It also adopts a rich and flexible application
  description format which is versatile enough to be used in various
  platforms. Boutiques is currently supported by \todo{X} platforms
  and was used to describe \todo{Y} applications.
}

@@

Platforms such as web services, portals, virtual research
environments, science gateways and workflow engines often rely on
third-party data processing and simulation applications to deliver
services to their users. Such applications are often repeatedly and
manually ported to various platforms whereas this porting effort would
greatly benefit from being mutualized. Virtual container systems such
as Docker and Singularity greatly facilitate the sharing of
applications executed in Linux systems, and improve the
reproducibility of analyses by defining immutable, reusable execution
environments. However, containers need to be properly annotated to
allow platforms to leverage the applications that they
contain. Basically, the command lines that can be run in a container
must be formally described so that they can be automatically created
and executed by platforms. Describing command lines is not
straightforward since they often contain implicit semantics specified
in human-readable documents such as manuals.

In Section~\ref{sec:system}, we describe \boutiques, a system to share
command-line applications across platforms. In \boutiques, a command
line is described from a flexible template referring to the inputs it
requires and the outputs it produces. Inputs may be inter-dependent,
e.g., input A is available only if input B was defined and passed in
configuration files or directly on the command line. Such formal
descriptions, also called \emph{manifests}, allows for advanced
validation of input values, which is useful error detection. Manifests
refer to containers where applications and all their dependencies are
installed. \boutiques also provides a simple parallelization framework
that allows applications such as workflow engines to execute multiple
tasks concurrently. A set of core tools facilitates the construction,
validation and execution of \boutiques manifests. A repository
orchestrates these tools together.

\boutiques was created using a bottom-up approach in which new
features were progressively added as applications and platforms
required. It was used to describe \todo{Y} applications and it is
supported by \todo{X} platforms, which we report in
Section~\ref{sec:results}.

\boutiques manifests are intented to be produced by scientific
application developers, to be published in common repositories, and to
be consumed by execution platforms. Although the motivating use cases
were taken from neuroinformatics, our primary field of interest,
nothing prevents the system from being used in a variety of other
fields.

\note{Tristan}{Vocabulary (check consistency across the paper): tool vs. application; descriptor vs. manifest; task vs. invocation}
\note{Tristan}{Targeted journals: Scientific Data, Giga Science, Frontiers in Neuroinformatics, Future Generation Computer Systems}

\section{System description}
\label{sec:system}

@@

\todo{json editor}

@@

\section{Results}
\label{sec:results}

@@

\subsection{Flywheel}

\todo{Gunnar, Nathanael, John}

\subsection{Pegasus}

@@

\subsection{SPINE}

@@

\subsection{VIP}
